MolProbity's Ultimate Rotamer-Library Distributions for Model Validation

### 2016-02-25 methods.tex

\subsection{Rotamer Assignment and Score}
\label{sec_md:rotaassign}
For the purpose of setting the appropriate contour level for the outlier cutoff, a large test dataset was created from all PDBs in the Top8000 by including all chains and all residues in each PDB file (as a surrogate for entire new PDB files later being validated), hereafter called the unfiltered dataset. \textit{phenix.rotalyze} was used to assign rotamer names and scores for each residue in the unfiltered dataset. The score is calculated from the individual residue's $\chi$ values, by interpolating over the nearest contour grid points (see section \ref{sec_md:contours}). This was done twice, once using the new Top8000 contours and once using our previous Top500 contours. If the score is above the outlier threshold, then a rotamer name is assigned based on which rotamer well the given $\chi$ angles fall within (the maximum well ranges were described in Section \ref{sec_md:chi_stats}). Otherwise the side-chain conformation is classified as an \textit{outlier}.

>>>>

\subsection{Rotamer Assignment and Score}
\label{sec_md:rotaassign}
For the purpose of setting the appropriate contour level for the new outlier cutoff, a large test dataset was created from all PDBs in the Top8000 by including all chains and all residues in each PDB file (as a surrogate for entire new PDB files later being validated), hereafter called the unfiltered dataset. \textit{phenix.rotalyze} was used to assign rotamer names and scores for each residue in the unfiltered dataset. The score is calculated from the individual residue's $\chi$ values, by interpolating over the nearest contour grid points (see section \ref{sec_md:contours}). This was done twice, once using the new Top8000 contours and once using our previous Top500 contours. If the score is above the outlier threshold, then a rotamer name is assigned based on which rotamer well the given $\chi$ angles fall within (the maximum well ranges were described in Section \ref{sec_md:chi_stats}). Otherwise the side-chain conformation is classified as an \textit{outlier.}  \textcolor{changecolor}{A Top8000 cutoff of 0.3\% flagged very nearly the same number of outliers in the unfiltered dataset as the current Top500 cutoff of 1.0\% , so the value of 0.3\% was chosen as the outlier contour level for the new system.}

\textcolor{changecolor}{The contour peaks for 214 rotamer clusters reach the 0.3\% level. To ensure validity, for clusters with $\le$ 8 non-outlier datapoints we manually examined model, electron density, and interactions for the example residues. Four such clusters were judged reliable and only two unreliable (both for Lys and with $\le$ 4 datapoints). There are therefore 212 named rotamers in the final list.}
